---
title: "Distance to centroid report"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)

## Set up the cache and config
db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analysis", "drake", "drake-cache.sqlite"))
cache <- storr::storr_dbi("datatable", "keystable", db)

```

```{r load stuff}
loadd(cd_long_df, cache = cache)
loadd(cds_df, cache = cache)
```

### Distribution of skewnesses
```{r distance violins, fig.height = 10, fig.width = 6} 

dist_violins <- ggplot(data = cds_df, aes(x = 0, y = centroid_dist)) +
  geom_violin(data = filter(cds_df, source == "sampled"),  aes(x = 0, y = centroid_dist, color = season)) +
  geom_point(data = filter(cds_df, source == "observed"),  aes(x = 0, y = centroid_dist, color = season)) +
  facet_grid(rows = vars(year), cols = vars(season), switch = "y") +
  scale_color_viridis_d(end = .8, direction = -1) +
  theme_bw()

dist_violins
```

2009 is problematic because the communities are very small (S = 1 and 3, N = 2 and 5 respectively). Similarly, winter of 2000 had 12 individuals of 2 species. Notably, winter 1996 had 34 individuals of 3 species, and this appears to be enough to get some interesting variation going.

So far, this result contrasts with what I was finding earlier. Skewness is not always an incredible outlier. I am going to do this with the 1994 data to confirm that I haven't messed up somewhere.

If this holds (and perhaps even if it doesn't) perhaps it's worth re-investigating Legendre approximation over a broader set of datasets. It looks like the ones I pulled might have been unusual. 

### Heatmaps 

```{r heatmap of fs, fig.height = 20, fig.width = 10}

ranked_cds_df <- cds_df %>%
  filter(source == "sampled") %>%
  group_by(year, season, treatment) %>%
  arrange(centroid_dist) %>%
  mutate(dist_rank = row_number()) %>%
  mutate(dist_percentile = 100*(dist_rank / max(dist_rank))) %>%
  ungroup() 

observed_percentile <- cds_df %>%
  filter(source == "observed")

observed_percentile$dist_percentile <- NA

for(i in 1:nrow(observed_percentile)) {
  this_pool <- filter(ranked_cds_df, year == observed_percentile$year[i],
                      season == observed_percentile$season[i])
  
  this_rank <-max(filter(this_pool, centroid_dist <= observed_percentile$centroid_dist[i])$dist_rank)
  
  observed_percentile$dist_percentile[i] = 100 * (this_rank / nrow(this_pool))
}

ranked_cds_df <- bind_rows(ranked_cds_df, observed_percentile)

ranked_cd_long <- cd_long_df %>%
  filter(year != 2009) %>%
  left_join(ranked_cds_df, by = c("sim", "source", "season", "year", "treatment"))

cd_labels <- ranked_cd_long %>%
  group_by(year, season, treatment) %>%
  mutate(rank_y = .75 * max(abund)) %>%
  ungroup() %>%
  filter(source == "observed") %>%
  group_by(year, season, treatment) %>%
  filter(rank == floor(median(rank))) %>%
  ungroup()

heatmaps <- ggplot(data = ranked_cd_long, aes(x = rank, y = abund, color = dist_percentile)) +
  geom_line(data = filter(ranked_cd_long, source == "sampled"), aes(x = rank, y = abund, color = dist_percentile, group = sim), alpha = .05) +
  geom_point(data = filter(ranked_cd_long, source == "observed"), aes(x = rank, y = abund, color = dist_percentile)) + 
  geom_label(data = cd_labels, aes(x = rank, y = rank_y, color = dist_percentile, label = paste0("Obs. %ile: ", signif(dist_percentile, digits = 3))), nudge_x = .25) +
  facet_wrap(facets = c("season", "year"), ncol = 2, strip.position = "top", scales = "free", dir = "v") +
  theme_bw() +
  scale_color_viridis_c(option = "plasma", end = .7)

heatmaps

```

### Relating observed skewness to S and N
```{r obs S and N, fig.height = 12, fig.width = 5}

obs_sn <- ranked_cd_long %>%
  filter(source == "observed") %>%
  group_by(season, year, treatment, dist_percentile) %>%
  summarize(nind = sum(abund),
            nspp = max(rank)) %>%
  ungroup() %>%
  mutate(year = as.factor(year))


dist_s <-ggplot(data = obs_sn, aes (x =nspp, y = dist_percentile, color = year)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_d(end = .7) +
  ggtitle("Centroid dist vs N / S")
  ggtitle("Skewness vs S")

dist_n <- ggplot(data = obs_sn, aes (x = nind, y = dist_percentile, color = year)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_d(end = .7) +
  ggtitle("Skewness vs N")

dist_ns <- ggplot(data = obs_sn, aes (x = nind / nspp, y = dist_percentile, color = year)) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_d(end = .7) +
  ggtitle("Skewness vs N / S")

gridExtra::grid.arrange(grobs = list(dist_s, dist_n, dist_ns), ncol = 1)
```
