---
title: "Diversity index report"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(drake)
library(dplyr)
library(ggplot2)
library(scadsplants)

# ## Set up the cache and config
# db <- DBI::dbConnect(RSQLite::SQLite(), here::here("analysis", "drake", "drake-cache.sqlite"))
# cache <- storr::storr_dbi("datatable", "keystable", db)

```

```{r load stuff}
di_df <- read.csv(here::here("analysis", "drake", "di_df.csv"), stringsAsFactors = F)
```

### Distribution of skewnesses
```{r dist of skew, fig.height = 10, fig.width = 6} 
di_df <- di_df %>%
  mutate(singletons = ifelse(grepl(di_name, pattern = "singles"), "singletons", "true"))

skew_violins <- ggplot(data = di_df, aes(x = singletons, y = skew, color = season)) +
  geom_violin(data = filter(di_df, source == "sampled")) +
  geom_point(data = filter(di_df, source == "observed")) +
 facet_wrap(c("year", "season"), ncol = 5, strip.position = "top") +
  scale_color_viridis_d(end = .8, direction = -1) +
  theme_bw()

skew_violins
```

2009 is problematic because the communities are very small (S = 1 and 3, N = 2 and 5 respectively). Similarly, winter of 2000 had 12 individuals of 2 species. Notably, winter 1996 had 34 individuals of 3 species, and this appears to be enough to get some interesting variation going.

So far, this result contrasts with what I was finding earlier. Skewness is not always an incredible outlier. I am going to do this with the 1994 data to confirm that I haven't messed up somewhere.

If this holds (and perhaps even if it doesn't) perhaps it's worth re-investigating Legendre approximation over a broader set of datasets. It looks like the ones I pulled might have been unusual. 


#### Shannon violins

```{r shannon violins,  fig.height = 10, fig.width = 6} 

shannon_violins <- ggplot(data = di_df, aes(x = singletons, y = shannon, color = season)) +
 geom_violin(data = filter(di_df, source == "sampled")) +
  geom_point(data = filter(di_df, source == "observed")) +
 facet_wrap(c("year", "season"), ncol = 5, strip.position = "top") +
  scale_color_viridis_d(end = .8, direction = -1) +
  theme_bw()

shannon_violins
```

#### Simpson violins

```{r simspon violins,  fig.height = 10, fig.width = 6} 

simpson_violins <- ggplot(data = di_df, aes(x = singletons, y = simpson, color = season)) +
 geom_violin(data = filter(di_df, source == "sampled")) +
  geom_point(data = filter(di_df, source == "observed")) +
 facet_wrap(c("year", "season"), ncol = 5, strip.position = "top") +
  scale_color_viridis_d(end = .8, direction = -1) +
  theme_bw()

simpson_violins
```

### Frequency of diversity index percentiles

```{r skewness percentile hist}

ranked_di_df <- di_df %>%
  filter(source == "sampled") %>%
  group_by(year, season, treatment, singletons) %>%
  mutate_at(c("skew", "shannon", "simpson"), .funs = list("percentile" = get_percentiles)) %>%
  ungroup()

observed_percentile <- di_df %>%
  filter(source == "observed") %>%
  left_join(select(ranked_di_df, -sim, -source, -skew_percentile, -shannon_percentile, -simpson_percentile, -di_name), by = c("year", "season", "treatment", "singletons")) %>%
  group_by(year, season, treatment, singletons) %>%
  mutate(skew_percentile = get_percentile(skew.x, a_vector = skew.y),
         shannon_percentile = get_percentile(shannon.x, a_vector = shannon.y),
         simpson_percentile = get_percentile(simpson.x, a_vector = simpson.y))  %>%
  ungroup() %>%
  select(year, season, treatment, source, singletons, skew_percentile, shannon_percentile, simpson_percentile,
         skew.x, shannon.x, simpson.x, di_name) %>%
  rename(skew = skew.x,
         shannon = shannon.x,
         simpson = simpson.x) %>%
  distinct()


ranked_di_df <- bind_rows(ranked_di_df, observed_percentile)

ranked_di_long <- ranked_di_df %>%
  tidyr::gather(-sim, -year, -season, -treatment, -source, -di_name, -singletons, key = "variable", value = "value")

di_hists <- ggplot(data = filter(ranked_di_long, source == "observed", singletons == "true", variable %in% c("skew_percentile", "shannon_percentile", "simpson_percentile")), aes(x = season, y= value,color = season)) +
  geom_violin() +
  geom_jitter(alpha = .7, height = 0, width = .2) +
  theme_bw() +
  scale_color_viridis_d(end = .8, direction = -1) +
  facet_wrap(~variable, strip.position = "top")

di_hists

save(ranked_di_df, ranked_di_long, file = here::here("analysis", "reports", "ranked_dis.Rds"))

```

### Sensitivity of percentile to adding singletons
```{r singletons 1to1, fig.width = 4.5, fig.height = 6}

singletons_1to1_dat <- filter(ranked_di_long, source == "observed", variable %in% c("skew_percentile", "shannon_percentile", "simpson_percentile"), singletons == "true") %>%
  select(-di_name) %>%
  left_join(filter(ranked_di_long, source == "observed", variable %in% c("skew_percentile", "shannon_percentile", "simpson_percentile"), singletons == "singletons"), by = c("sim", "year", "season", "treatment", "source", "variable")) %>%
  select(-sim, -di_name) %>%
  rename(true_value = value.x,
         singletons_value = value.y) %>%
  select(-singletons.x, -singletons.y)

singletons_1to1_plot <- ggplot(data = singletons_1to1_dat, aes(x = true_value, y = singletons_value, color = season)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, alpha = .5) +
  theme_bw() +
  facet_wrap(vars(variable, season), ncol = 2) +  
  scale_color_viridis_d(end = .8, direction = -1)


singletons_1to1_plot
```


